Aligning marine species range data to better serve science and conservation

### 2016-02-17 data_explore/data_prep.Rmd
Updating numbers etc in manuscript... done with "data poor" section?

library(raster)
library(readr)
library(data.table)
library(ggplot2)
library(tidyr)
library(dplyr)
library(stringr)

@@

if(basename(getwd()) != 'data_explore') setwd('~/github/IUCN-AquaMaps/data_explore')
source('data_explore_fxns.R')
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

@@

``` {r pull_files_to_local, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
### pull spp_iucn_marine.csv from git-annex and bring to local repo
spp_iucn <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_cleaned.csv'), 
                            stringsAsFactors = FALSE)
spp_iucn <- spp_iucn %>%
  dplyr::select(sciname, iucn_sid, iucn_category, 
                popn_trend, parent_sid, subpop_sid) %>%
  unique()
write_csv(spp_iucn, file.path(dir_data, 'spp_iucn.csv'))

### pull spp_iucn_maps_all.csv from git-annex and bring to local repo
spp_iucn_maps <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_maps_all.csv'), 
                            stringsAsFactors = FALSE) %>%
  rename(iucn_sid = id_no)

@@

### pull cleaned spp_am data from git-annex and bring to local repo
spp_am <- read.csv(file.path(dir_anx, scenario, 'int/spp_am_cleaned.csv'), 
                   stringsAsFactors = FALSE)
write_csv(spp_am, file.path(dir_data, 'spp_am.csv'))

### pull raw aquamaps species list from git-annex and bring to local repo
spp_am_raw <- read.csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv'), 
                   stringsAsFactors = FALSE)
write_csv(spp_am_raw, file.path(dir_data, 'speciesoccursum.csv'))

@@

```{r get_spp_all, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
### get spp_all_complete from git-annex, bring to local repo.  The spp_all_complete
### file was created in ohiprep/globalprep/SPP_ICO/data_prep_SPP.R, running the
### function create_spp_master_lookup(source_pref = 'iucn', fn_tag = '_complete', reload = TRUE)
spp_all_remote <- read.csv(file.path(dir_anx, scenario, 'int/spp_all.csv'), 
                           stringsAsFactors = FALSE) 
spp_all_remote_clean <- spp_all_remote %>%
  filter(!str_detect(spatial_source, 'subpop')) %>%
  filter(!str_detect(spatial_source,  'alias'))
### At this point, plenty of duplicates:
### * This code returns 984 duplicated names:
###     length(spp_all_remote_clean$sciname) - length(unique(spp_all_remote_clean$sciname))
### * some are AM subpops which were ignored in processing the remote spp_all file.
###   Note: All subpops have an identified parent, so can filter them out without
###   losing any species.
###   code to verify this:
###     z_parent <- spp_all_remote_clean %>% 
###       filter(!is.na(subpop_sid) & spatial_source == 'am')
###     z_subpop <- spp_all_remote_clean %>% 
###       filter(!is.na(parent_sid) & spatial_source == 'am')
###     nrow(z_subpop %>% filter(!sciname %in% z_parent$sciname))
### * some are duplicated IUCN parents (e.g. Dermochelys coriacea) that will be 
###   eliminated when select(-iucn_subpop) column and then unique()

@@

### But, some of the extracted IUCN maps don't appear on this list! add 'em in.
spp_iucn_maps_new <- read.csv(file.path(dir_data, 'spp_iucn_maps.csv'), 
                              stringsAsFactors = FALSE) %>%
  filter(!sciname %in% spp_all$sciname) %>%
  filter(!str_detect(sciname, 'subpop')) %>% ### a Hawaiian subpop... ditch it!
  select(-iucn_subpop)

@@

### Also, some of the AM listings have disappeared due to name conflicts
### - aliases, etc.
spp_am_new <- read.csv(file.path(dir_data, 'speciesoccursum.csv'), 
                       stringsAsFactors = FALSE) %>%
  mutate(sciname = paste(str_trim(genus), str_trim(species)),
         iucn_sid = as.integer(iucn_id)) %>%
  dplyr::select(am_sid = speciesid, sciname, reviewed, iucn_sid,
                popn_category = iucn_code) %>%
  filter(!am_sid %in% spp_all$am_sid) %>%
  mutate(spatial_source = 'am',
         popn_category = ifelse(popn_category == 'LR/lc', 'LC', popn_category),
         popn_category = ifelse(popn_category == 'LR/nt', 'NT', popn_category),
         popn_category = ifelse(popn_category == 'N.E.',  'NE', popn_category))

@@

```{r get_other_vars, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
### create spp_am_v_iucn: the whole maplist narrowed down to just the overlapping species.
### Overlaps detected from spatial_source containing 'iucn' (b/c source_pref = 'iucn' above),
### that also have a non-NA AquaMaps ID code.
spp_all <- read.csv(file.path(dir_data, 'spp_all.csv'), stringsAsFactors = FALSE)

@@

loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
loiczid_raster       <- raster(loiczid_raster_file)
names(loiczid_raster) <- 'loiczid'

@@

data_file <- file.path(dir_data, 'spp_list_w_area_all.csv')
reload <- TRUE
if (!file.exists(data_file) | reload) {
  ### species info list - scinames, IUCN species ID, AM species ID, category etc
  spp_list_base <- read_csv(file.path(dir_data, 'spp_am_v_iucn.csv'))

@@

spp_list <- read.csv(data_file, stringsAsFactors = FALSE) 
### At this point, spp_list zero-area IUCNs due to missing (?) polygons (maybe points?)

@@

### attach group name text field, and occurcells field
spp_list <- spp_list %>%
  left_join(read.csv(file.path(dir_data, 'sciname_lookups/spp_groupnames.csv'), 
                     stringsAsFactors = FALSE),
            by = 'spp_group') %>%
  left_join(read.csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv'), 
                   stringsAsFactors = FALSE) %>% 
              select(am_sid = speciesid, occurcells),
            by = 'am_sid')

@@

``` {r counts_spp_all_observations}
spp_all <- read_csv(file.path(dir_data, 'spp_all.csv'))

library(rgbif)

@@

# write_csv(gbif_occur, file.path(dir_data, 'gbif_occurrence_df.csv'))
gbif_occur <- read_csv(file.path(dir_data, 'gbif_occurrence_df.csv'))

source('obis_count.R')
scinames <- spp_all$sciname

obis_occur <- obis_count(scinames)

obis_occur <- obis_occur %>% 
  left_join(spp_all %>%
              select(iucn_sid, am_sid, sciname),
            by = 'sciname') %>%
  left_join(read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv')) %>%
                       select(am_sid = speciesid, occurcells),
                     by = 'am_sid')

write_csv(obis_occur, file.path(dir_data, 'obis_occurrence.csv'))

totes_occur <- obis_occur %>%
  left_join(gbif_occur)

@@

>>>>

library(readr)
library(data.table)
library(tidyr)
library(dplyr)
library(stringr)

@@

if(basename(getwd()) != 'data_explore') setwd('~/github/IUCN-AquaMaps/data_explore')
# source('data_explore_fxns.R')
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

@@

``` {r pull_files_to_local, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
### pull spp_iucn_marine.csv from git-annex and bring to local repo
spp_iucn <- read_csv(file.path(dir_anx, scenario, 'int/spp_iucn_cleaned.csv'))
spp_iucn <- spp_iucn %>%
  dplyr::select(sciname, iucn_sid, iucn_category, 
                popn_trend, parent_sid, subpop_sid) %>%
  unique()
write_csv(spp_iucn, file.path(dir_data, 'spp_iucn.csv'))

### pull spp_iucn_maps_all.csv from git-annex and bring to local repo
spp_iucn_maps <- read_csv(file.path(dir_anx, scenario, 'int/spp_iucn_maps_all.csv')) %>%
  rename(iucn_sid = id_no)

@@

### pull cleaned spp_am data from git-annex and bring to local repo
spp_am <- read_csv(file.path(dir_anx, scenario, 'int/spp_am_cleaned.csv'))
write_csv(spp_am, file.path(dir_data, 'spp_am.csv'))

### pull raw aquamaps species list from git-annex and bring to local repo
spp_am_raw <- read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv'))
write_csv(spp_am_raw, file.path(dir_data, 'speciesoccursum.csv'))

@@

```{r get_spp_all, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
### get spp_all_complete from git-annex, bring to local repo.  The spp_all_complete
### file was created in ohiprep/globalprep/SPP_ICO/data_prep_SPP.R, running the
### function create_spp_master_lookup(source_pref = 'iucn', fn_tag = '_complete', reload = TRUE)
spp_all_remote <- read_csv(file.path(dir_anx, scenario, 'int/spp_all.csv'))
spp_all_remote_clean <- spp_all_remote %>%
  filter(!str_detect(spatial_source, 'subpop')) %>%
  filter(!str_detect(spatial_source,  'alias'))
### At this point, plenty of duplicates:
### * This code returns 984 duplicated names:
###     length(spp_all_remote_clean$sciname) - length(unique(spp_all_remote_clean$sciname))
### * some are AM subpops which were ignored in processing the remote spp_all file.
###   Note: All subpops have an identified parent, so can filter them out without
###   losing any species.
###   code to verify this:
###     z_parent <- spp_all_remote_clean %>% 
###       filter(!is.na(subpop_sid) & spatial_source == 'am')
###     z_subpop <- spp_all_remote_clean %>% 
###       filter(!is.na(parent_sid) & spatial_source == 'am')
###     nrow(z_subpop %>% filter(!sciname %in% z_parent$sciname))
### * some are duplicated IUCN parents (e.g. Dermochelys coriacea) that will be 
###   eliminated when select(-iucn_subpop) column and then unique()

@@

### But, some of the extracted IUCN maps don't appear on this list! add 'em in.
spp_iucn_maps_new <- read_csv(file.path(dir_data, 'spp_iucn_maps.csv')) %>%
  filter(!sciname %in% spp_all$sciname) %>%
  filter(!str_detect(sciname, 'subpop')) %>% ### a Hawaiian subpop... ditch it!
  select(-iucn_subpop)

@@

### Also, some of the AM listings have disappeared due to name conflicts
### - aliases, etc.
spp_am_new <- read_csv(file.path(dir_data, 'speciesoccursum.csv')) %>%
  mutate(sciname = paste(str_trim(genus), str_trim(species)),
         iucn_sid = as.integer(iucn_id)) %>%
  dplyr::select(am_sid = speciesid, sciname, reviewed, iucn_sid,
                popn_category = iucn_code) %>%
  filter(!am_sid %in% spp_all$am_sid) %>%
  mutate(spatial_source = 'am',
         popn_category = ifelse(popn_category == 'LR/lc', 'LC', popn_category),
         popn_category = ifelse(popn_category == 'LR/nt', 'NT', popn_category),
         popn_category = ifelse(popn_category == 'N.E.',  'NE', popn_category))

@@

```{r get_other_vars, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
### create spp_am_v_iucn: the whole maplist narrowed down to just the overlapping species.
### Overlaps detected from spatial_source containing 'iucn' (b/c source_pref = 'iucn' above),
### that also have a non-NA AquaMaps ID code.
spp_all <- read_csv(file.path(dir_data, 'spp_all.csv'))

@@

@@

loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
loiczid_raster       <- raster::raster(loiczid_raster_file)
names(loiczid_raster) <- 'loiczid'

data_file <- file.path(dir_data, 'spp_list_w_area_all.csv')
reload <- FALSE
if (!file.exists(data_file) | reload) {
  ### species info list - scinames, IUCN species ID, AM species ID, category etc
  spp_list_base <- read_csv(file.path(dir_data, 'spp_am_v_iucn.csv'))

@@

spp_list <- read_csv(data_file) 
### At this point, spp_list zero-area IUCNs due to missing (?) polygons (maybe points?)

@@

### attach group name text field, and occurcells field
spp_list <- spp_list %>%
  left_join(x <- read_csv(file.path(dir_data, 'sciname_lookups/spp_groupnames.csv')),
            by = 'spp_group') %>%
  left_join(read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv')) %>% 
              select(am_sid = speciesid, occurcells),
            by = 'am_sid')

@@

``` {r counts_spp_all_observations, echo = FALSE}
spp_all <- read_csv(file.path(dir_data, 'spp_all.csv'))

# library(rgbif)

@@

# write_csv(gbif_occur, file.path(dir_data, 'gbif_occurrence.csv'))
gbif_occur <- read_csv(file.path(dir_data, 'gbif_occurrence.csv'))

# source('obis_count.R')
# scinames <- spp_all$sciname
# 
# obis_occur <- obis_count(scinames)
# obis_occur <- obis_count(scinames[26500:26600], verbose = TRUE)
# 
# obis_occur <- obis_occur %>% 
#   left_join(spp_all %>%
#               select(iucn_sid, am_sid, sciname),
#             by = 'sciname') %>%
#   left_join(read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv')) %>%
#                        select(am_sid = speciesid, occurcells),
#                      by = 'am_sid')
# 
# write_csv(obis_occur, file.path(dir_data, 'obis_occurrence.csv'))
obis_occur <- read_csv(file.path(dir_data, 'obis_occurrence.csv'))

totes_occur <- obis_occur %>%
  left_join(gbif_occur) %>%
  rowwise() %>%
  mutate(mean_occ = mean(c(gbif_occ, obis_occ), na.rm = TRUE))

@@

__Possible "data poor" indicators:__

* (of matched species) how does bottom decile get distributed across quadrants
* use mean of OBIS and GBIF, and just report average (or median?) occurrences across all species in each quadrant
* also use similar method for AM occurcells

``` {r data_poor_analysis, echo = FALSE, eval = FALSE}
spp_quads <- read_csv(file.path(dir_data, 'spp_list_quads.csv')) %>% 
  select(am_sid, iucn_sid, occurcells, quad) %>%
  unique()
### This contains all mapped species, discarding those with no actual IUCN maps

spp_occ_counts <- spp_quads %>%
  left_join(totes_occur) %>%
  group_by(iucn_sid, am_sid, quad) %>%
  summarize(occurcells = mean(occurcells, na.rm = TRUE), occ = mean(mean_occ, na.rm = TRUE)) %>%
  group_by(iucn_sid, quad) %>%
  summarize(occurcells = sum(occurcells, na.rm = TRUE), occ = sum(occ, na.rm = TRUE)) %>%
  ungroup()

btm_dec_occurcells <- spp_occ_counts$occurcells %>% quantile(c(.1, .25))
btm_dec_occ <- spp_occ_counts$occ %>% quantile(c(.1, .25))

quad_count_means <- spp_occ_counts %>%
  mutate(quad = ifelse(quad == 'q4', 'q4', 'q123')) %>%
  group_by(quad) %>%
  summarize(mean_occurcells = mean(occurcells),
            mean_occ        = mean(occ),
            med_occurcells  = median(occurcells),
            med_occ         = median(occ),
            dp_am_10 = sum(occurcells <= btm_dec_occurcells[1])/n(),
            dp_am_25 = sum(occurcells <= btm_dec_occurcells[2])/n(),
            dp_gb_10 = sum(occ <= btm_dec_occ[1])/n(),
            dp_gb_25 = sum(occ <= btm_dec_occ[2])/n())

write_csv(quad_count_means, file.path(dir_data, 'quads_data_poor.csv'))

### 2016-02-17 manuscript-main.Rmd
Updating numbers etc in manuscript... done with "data poor" section?

# library(ggplot2)
# library(maps)
# library(readr)

# library(raster)
# library(data.table)
library(tidyr)
library(dplyr)
library(stringr)

@@

# dir_anx  <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
# dir_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data')
# dir_fig  <- file.path(getwd(), 'figures')

@@

@@

data_file <- 'data/spp_list_w_area_trimmed.csv'
spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE)
spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) 
### NOTE: as of Jan 31, includes all map pairs, even ones where IUCN area == 0

### generate other lists
spp_all_file <- 'data/spp_all.csv'
spp_all <- read.csv(spp_all_file, stringsAsFactors = FALSE)

# iucn_marine_file <- 'data/spp_iucn_marine.csv'
# spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)

iucn_maps_file <- 'data/spp_iucn_maps.csv'
spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE)

spp_am_file <- 'data/spp_am.csv'
spp_am <- read.csv(spp_am_file, stringsAsFactors = FALSE)

>>>>

# library(ggplot2)
# library(maps)

# library(raster)
# library(data.table)
library(tidyr)
library(dplyr)
library(stringr)
library(readr)

@@

# dir_anx  <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')

@@

dir_data <- 'data'
dir_fig  <- 'figures'

@@

data_file <- 'data/spp_list_w_area_trimmed.csv'
spp_map_pairs <- read_csv(data_file)
### NOTE: as of Jan 31, includes all map pairs, even ones where IUCN area == 0

### generate other lists
spp_all_file <- 'data/spp_all.csv'
spp_all <- read_csv(spp_all_file)

# iucn_marine_file <- 'data/spp_iucn_marine.csv'
# spp_iucn_marine <- read_csv(iucn_marine_file)

iucn_maps_file <- 'data/spp_iucn_maps.csv'
spp_iucn_maps <- read_csv(iucn_maps_file)

spp_am_file <- 'data/spp_am.csv'
spp_am <- read_csv(spp_am_file)

### 2016-02-17 ms_2_intro_res_disc.Rmd
committing so I can re-run area calcs
Updating numbers etc in manuscript... done with "data poor" section?

``` {r calc species numbers, echo = FALSE}
n_spp_all    <- nrow(spp_all)
n_spp_am     <- nrow(spp_am)
n_spp_am_rev <- nrow(spp_am %>% filter(reviewed == 1))
n_spp_iucn   <- nrow(spp_iucn_maps)
n_spp_both   <- nrow(spp_map_pairs)

@@

To understand the implications of differences between the AquaMaps and IUCN datasets, we compare how each data source represents the global spatial and taxonomic distribution of species.  Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` species).  For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align, determine several issues that lead to misalignment between predicted species distributions, and outline possible improvements. <!--BSH: I think Jamie makes this comment below, but we may need to be able to substantiate this somehow.  Is there a way to quickly figure out how many studies have used AM vs. IUCN?  CCO: What about age - AquaMaps is pretty recent; before that, would IUCN have been the dominant dataset for marine?  Maybe we can track citations of AquaMaps, but citations of IUCN I think are generic - just cite the Red List - not referencing spatial information specifically, much less marine.-->

@@

n_q4_rev <- nrow(spp_list_quads %>% filter(quad == 'q4' & !is.na(reviewed)))
p_q4_rev <- round(n_q4_rev/n_q4 * 100, 1)
n_q123_rev <- nrow(spp_list_quads %>% filter(quad != 'q4' &  !is.na(reviewed)))
p_q123_rev <- round(n_q123_rev/n_spp_both * 100, 1)
```

@@

<!--The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor extent match.  Often these mismatches arise when AquaMaps relies on limited data to predict species presence well beyond known occurrences.  Indeed, `r p_q4_datapoor`% (n = `r n_q4_datapoor`) of these cases are data-poor (in which the AquaMaps environmental envelope model is based on fewer than 10 known cell occurrences), compared with `r p_q123_datapoor`% for species in the other three quadrants.  Only `r p_q4_rev`% (n = `r n_q4_rev`) of AquaMaps maps in this quadrant have undergone expert review, compared with `r p_q123_rev`% of species in the other three quadrants. <!--BSH: This implies the problem is solely with AM.  Are there no issues with IUCN maps that cause problems in this quadrant?--> ___The data-poor condition likely applies to IUCN as well - AM using OBIS points to generate models, and IUCN is probably doing the same.  Check, and rework.  This means AM is creating errors of commission, while IUCN may be committing errors of omission for poorly-studied species and/or regions - compounding the total problem.___-->

The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor extent match.  When extrapolating from limited observations, the AquaMaps model often predicts species presence well beyond known occurrences; at the same time, IUCN range maps .  Indeed, `r p_q4_datapoor`% (n = `r n_q4_datapoor`) of these cases are data-poor (in which the AquaMaps environmental envelope model is based on fewer than 10 known cell occurrences ___is there a good way to get the number of OBIS observations for each of these species, as an external validation of data-poor?)___, compared with `r p_q123_datapoor`% for species in the other three quadrants.  Only `r p_q4_rev`% (n = `r n_q4_rev`) of AquaMaps maps in this quadrant have undergone expert review, compared with `r p_q123_rev`% of species in the other three quadrants. <!--BSH: This implies the problem is solely with AM.  Are there no issues with IUCN maps that cause problems in this quadrant?--> ___The data-poor condition likely applies to IUCN as well - AM using OBIS points to generate models, and IUCN is probably doing the same.  Check, and rework.  This means AM is creating errors of commission, while IUCN may be committing errors of omission for poorly-studied species and/or regions - compounding the total problem.___

>>>>

``` {r calc species numbers, echo = FALSE}
n_spp_all    <- nrow(spp_all)
n_spp_am     <- nrow(spp_am)
n_spp_am_rev <- nrow(spp_am %>% filter(reviewed == 1))
n_spp_iucn   <- length(spp_iucn_maps$iucn_sid %>% unique())
n_spp_both   <- length(spp_map_pairs$iucn_sid %>% unique())

@@

To understand the implications of differences between the AquaMaps and IUCN datasets, we compare how each data source represents the global spatial and taxonomic distribution of species.  Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` unique species).  For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align, determine several issues that lead to misalignment between predicted species distributions, and outline possible improvements. <!--BSH: I think Jamie makes this comment below, but we may need to be able to substantiate this somehow.  Is there a way to quickly figure out how many studies have used AM vs. IUCN?  CCO: What about age - AquaMaps is pretty recent; before that, would IUCN have been the dominant dataset for marine?  Maybe we can track citations of AquaMaps, but citations of IUCN I think are generic - just cite the Red List - not referencing spatial information specifically, much less marine.-->

@@

n_q4_rev <- nrow(spp_list_quads %>% filter(quad == 'q4' & !is.na(reviewed)))
p_q4_rev <- round(n_q4_rev/n_q4 * 100, 1)
n_q123_rev <- nrow(spp_list_quads %>% filter(quad != 'q4' &  !is.na(reviewed)))
p_q123_rev <- round(n_q123_rev/n_spp_both * 100, 1)

quad_count_means <- read_csv(file.path(dir_data, 'quads_data_poor.csv')) %>%
  arrange(quad)
dp_med_am_q4    <- quad_count_means$med_occurcells[4]
dp_med_am_q123  <- quad_count_means$med_occurcells_total[1]
dp_med_occ_q4   <- quad_count_means$med_occ[4]
dp_med_occ_q123 <- quad_count_means$med_occ_total[1]

```

@@

<!--The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor extent match.  Often these mismatches arise when AquaMaps relies on limited data to predict species presence well beyond known occurrences.  Indeed, `r p_q4_datapoor`% (n = `r n_q4_datapoor`) of these cases are data-poor (in which the AquaMaps environmental envelope model is based on fewer than 10 known cell occurrences), compared with `r p_q123_datapoor`% for species in the other three quadrants.  Only `r p_q4_rev`% (n = `r n_q4_rev`) of AquaMaps maps in this quadrant have undergone expert review, compared with `r p_q123_rev`% of species in the other three quadrants. <!--BSH: This implies the problem is solely with AM.  Are there no issues with IUCN maps that cause problems in this quadrant?--> <!--___The data-poor condition likely applies to IUCN as well - AM using OBIS points to generate models, and IUCN is probably doing the same.  Check, and rework.  This means AM is creating errors of commission, while IUCN may be committing errors of omission for poorly-studied species and/or regions - compounding the total problem.___-->

The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor extent match well. Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (summing (averaging?) OBIS and GBIF occurrences) for this quadrant is `r dp_med_am_q4`, compared to a median of `r dp_med_am_q123` for species in the other three quadrants.  When extrapolating from limited observations, the AquaMaps model often predicts species presence well beyond known occurrences, introducing commission errors; at the same time, IUCN range maps generally target known occurrences, possibly introducing omission errors for data-limited species.
